---
title: "financial_eco_MHP"
author: "Marc-Henri Pélet"
date: "24/05/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE, fig.width=10, fig.height=5)
options(width=120)
library(lattice)
library(ggplot2)
library(plyr)
library(randomForest)
```

Dataset : [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har) 


## Processing

```{r}
trainer.raw <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"))
tester.raw <- read.csv(url("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"))
```

## Data processing : the beginning

Look at the dimensions & head of the dataset to get an idea
```{r}
# Res 1
dim(trainer.raw)
# Res 2 - excluded because over the required amount
# head(trainer.raw)
# Res 3 - excluded because over the required amount
#str(trainer.raw)
# Res 4 - excluded because over the required amount
#summary(trainer.raw)
```


We remove the empty data 

```{r}
maxNAPercentage = 20
maxNACount <- nrow(trainer.raw) / 100 * maxNAPercentage
removeColumns <- which(colSums(is.na(trainer.raw) | trainer.raw=="") > maxNACount)
trainer.cleaned01 <- trainer.raw[,-removeColumns]
tester.cleaned01 <- tester.raw[,-removeColumns]
```

As we do not need the related data, we take out these too

```{r}
removeColumns <- grep("timestamp", names(trainer.cleaned01))
trainer.cleaned02 <- trainer.cleaned01[,-c(1, removeColumns )]
tester.cleaned02 <- tester.cleaned01[,-c(1, removeColumns )]
```


```{r}
classeLevels <- levels(trainer.cleaned02$classe)
trainer.cleaned03 <- data.frame(data.matrix(trainer.cleaned02))
trainer.cleaned03$classe <- factor(trainer.cleaned03$classe, labels=classeLevels)
tester.cleaned03 <- data.frame(data.matrix(tester.cleaned02))
```

We have now the final datas

```{r}
trainer.cleaned <- trainer.cleaned03
tester.cleaned <- tester.cleaned03
```


## Analysis


```{r}
set.seed(15691997)
library(caret)
clref <- which(names(trainer.cleaned) == "classe")
partition <- createDataPartition(y=trainer.cleaned$classe, p=0.7, list=FALSE)
trainer.subSetTrain <- trainer.cleaned[partition, ]
trainer.subSetTest <- trainer.cleaned[-partition, ]
```


```{r}
correlations <- cor(trainer.subSetTrain[, -clref], as.numeric(trainer.subSetTrain$classe))
optimalcorr <- subset(as.data.frame(as.table(correlations)), abs(Freq)>0.3)
optimalcorr
```

Normally, the best correlations are not above 0.3

We now plot this in order to have a better visualisation

```{r}
library(Rmisc)
library(ggplot2)
p1 <- ggplot(trainer.subSetTrain, aes(classe,pitch_forearm)) + 
  geom_boxplot(aes(fill=classe))
p2 <- ggplot(trainer.subSetTrain, aes(classe, magnet_arm_x)) + 
  geom_boxplot(aes(fill=classe))
multiplot(p1,p2,cols=2)
```


## Models 

Let's identify variables with high correlations amongst each other

We will then check if these modifications make the model more accurate 

```{r}
library(corrplot)
cormat <- cor(trainer.subSetTrain[, -clref])
highcor <- findCorrelation(cormat, cutoff=0.9, exact=TRUE)
excludeColumns <- c(highcor, clref)
corrplot(cormat, method="color", type="lower", order="hclust", tl.cex=0.70, tl.col="red", tl.srt = 45, diag = FALSE)
```

Therefore, some data seem corrolated with each other. We should then exclude them in order to have a better hindsight on the models.

```{r}
pcaPreProcess.all <- preProcess(trainer.subSetTrain[, -clref], method = "pca", thresh = 0.99)
trainer.subSetTrain.pca.all <- predict(pcaPreProcess.all, trainer.subSetTrain[, -clref])
trainer.subSetTest.pca.all <- predict(pcaPreProcess.all, trainer.subSetTest[, -clref])
tester.pca.all <- predict(pcaPreProcess.all, tester.cleaned[, -clref])
pcaPreProcess.subset <- preProcess(trainer.subSetTrain[, -excludeColumns], method = "pca", thresh = 0.99)
trainer.subSetTrain.pca.subset <- predict(pcaPreProcess.subset, trainer.subSetTrain[, -excludeColumns])
trainer.subSetTest.pca.subset <- predict(pcaPreProcess.subset, trainer.subSetTest[, -excludeColumns])
tester.pca.subset <- predict(pcaPreProcess.subset, tester.cleaned[, -clref])
```

Now we use the Random Forest trainer model with 200 trees

```{r}
library(randomForest)
ntree <- 150  
start <- proc.time()
rfMod.cleaned <- randomForest(
  x=trainer.subSetTrain[, -clref], 
  y=trainer.subSetTrain$classe,
  xtest=trainer.subSetTest[, -clref], 
  ytest=trainer.subSetTest$classe, 
  ntree=ntree,
  keep.forest=TRUE,
  proximity=TRUE) 
proc.time() - start
start <- proc.time()
rfMod.exclude <- randomForest(
  x=trainer.subSetTrain[, -excludeColumns], 
  y=trainer.subSetTrain$classe,
  xtest=trainer.subSetTest[, -excludeColumns], 
  ytest=trainer.subSetTest$classe, 
  ntree=ntree,
  keep.forest=TRUE,
  proximity=TRUE) 
proc.time() - start
start <- proc.time()
rfMod.pca.all <- randomForest(
  x=trainer.subSetTrain.pca.all, 
  y=trainer.subSetTrain$classe,
  xtest=trainer.subSetTest.pca.all, 
  ytest=trainer.subSetTest$classe, 
  ntree=ntree,
  keep.forest=TRUE,
  proximity=TRUE) 
proc.time() - start
start <- proc.time()
rfMod.pca.subset <- randomForest(
  x=trainer.subSetTrain.pca.subset, 
  y=trainer.subSetTrain$classe,
  xtest=trainer.subSetTest.pca.subset, 
  ytest=trainer.subSetTest$classe, 
  ntree=ntree,
  keep.forest=TRUE,
  proximity=TRUE) 
proc.time() - start
```

#Model examination

We will check the accuracies of each of the 4 models

```{r}
rfMod.cleaned
rfMod.cleaned.trainer.accuracy <- round(1-sum(rfMod.cleaned$confusion[, 'class.error']),3)
paste0("Accuracy trainer: ",rfMod.cleaned.trainer.accuracy)
rfMod.cleaned.tester.accuracy <- round(1-sum(rfMod.cleaned$test$confusion[, 'class.error']),3)
paste0("Accuracy tester: ",rfMod.cleaned.tester.accuracy)
rfMod.exclude
rfMod.exclude.trainer.accuracy <- round(1-sum(rfMod.exclude$confusion[, 'class.error']),3)
paste0("Accuracy trainer: ",rfMod.exclude.trainer.accuracy)
rfMod.exclude.tester.accuracy <- round(1-sum(rfMod.exclude$test$confusion[, 'class.error']),3)
paste0("Accuracy tester: ",rfMod.exclude.tester.accuracy)
rfMod.pca.all
rfMod.pca.all.trainer.accuracy <- round(1-sum(rfMod.pca.all$confusion[, 'class.error']),3)
paste0("Accuracy trainer: ",rfMod.pca.all.trainer.accuracy)
rfMod.pca.all.tester.accuracy <- round(1-sum(rfMod.pca.all$test$confusion[, 'class.error']),3)
paste0("Accuracy tester: ",rfMod.pca.all.tester.accuracy)
rfMod.pca.subset
rfMod.pca.subset.trainer.accuracy <- round(1-sum(rfMod.pca.subset$confusion[, 'class.error']),3)
paste0("Accuracy trainer: ",rfMod.pca.subset.trainer.accuracy)
rfMod.pca.subset.tester.accuracy <- round(1-sum(rfMod.pca.subset$test$confusion[, 'class.error']),3)
paste0("Accuracy tester: ",rfMod.pca.subset.tester.accuracy)
```

#Conclusion

The `rfMod.exclude` performs better then the 'rfMod.cleaned'

We will thereofre choose the  `rfMod.exclude` model as the best model to use for predicting the test set as it has the higher accuracy and the lowest error rate


We will now plot this model

```{r}
par(mfrow=c(1,2)) 
varImpPlot(rfMod.exclude, cex=0.6, pch=20, main='Variable Importance: rfMod.exclude')
plot(rfMod.exclude, cex=0.6, main='Error compared to number of trees')
par(mfrow=c(1,1)) 
```

#Results

We will run all four models for this final test

```{r}
predictions <- t(cbind(
    exclude=as.data.frame(predict(rfMod.exclude, tester.cleaned[, -excludeColumns]), optional=TRUE),
    cleaned=as.data.frame(predict(rfMod.cleaned, tester.cleaned), optional=TRUE),
    pcaAll=as.data.frame(predict(rfMod.pca.all, tester.pca.all), optional=TRUE),
    pcaExclude=as.data.frame(predict(rfMod.pca.subset, tester.pca.subset), optional=TRUE)
))
predictions
```

As we can see, there are not a lot of change between the results in these models. However, due to better accuracy, and mostly lower error rate, we will stick with the `rfMod.exclude` model